---
title: "README"
output: github_document
---

# anon

A library containing data anonymization tools for R.

## 1. Installation

```{}
# 1. Install "devtools" package on R
    install.packages("devtools")

# 2. Install the "anon" package on R
    devtools::install_github("rajalah71/anon") # or /anon@paketti
```

```{r}
# 3. Load the "anon" library
    library(anon)
```

## 2. Usage

The package contains 4 functions to anonymize your data: $k$-anonymity, $l$-diversity, RSA encryption and spectral anonymization with 2 helper functions provided. 

The package also contains functions to measure the success of the anonymization on three levels: privacy, utility and similarity. 

### 2.1 _k_-anonymity

```{r, results='hide', message=FALSE, warning=FALSE}
data("iris")

kanon = kAnon(iris, k=5, quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species"))[1:10, ]
```

```{r}
kanon
```

- This uses all columns as quasi identifiers by default and operates on them with mean on numeric columns and with mode on categorical ones.

- You can define your own generalization functions and on what columns the function will operate on. These functions must be on a list with item names matching the names of the quasi-identifier columns. The columns will be worked on the order you name them, or by cardinality on descending order if left unnamed. 

### 2.2 _l_-diversity

```{r, results='hide', message=FALSE, warning=FALSE}
ldiv = lDiversity(iris, sensitiveAttributes = "Species", quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"), l=2)[1:10, ] 
```

```{r}
ldiv
```

- Works similiarly to kAnon, you just need to specify the sensitive attribute(s). If your data has multiple sensitive attributes with too high cardinality, one can use the 'sensitive_generalizer' function to reduce the cardinality of those variables. This, however, will negatively impact the utility of the data.

### 2.3 RSA

```{r, results='hide', message=FALSE, warning=FALSE}
rsa = encrypt(iris)[1:10, ]
```

```{r}
rsa
```

- Encrypts the data using a new RSA key = 2048 bits by default. Do not distribute the public key.

### 2.4 Spectral anonymization

```{}
# General usage:
spectral(iris, your_anon_function, on_matrices = "U")
```

- Anonymizes data on the spectral basis provided by SVD, $M = UDV'$. Supports any anonymization function $f: R^{a \times b} \to R^{a \times b}$. In this package, helper functions 'cell_swap' and 'sensitive_noise' are provided for permutations and additive noise respectively.

```{r, results='hide', message=FALSE, warning=FALSE}
# Example usage, spectral permutation:
# Performs column permutations on each column of the matrix U
swap = spectral(iris, cell_swap, on_matrices = "U")
```

```{r}
swap[1:10, ]
```

```{r, results='hide', message=FALSE, warning=FALSE}
# Example usage spectral noise:
# Adds random noise on the columns of the matrix UD
noise = spectral(iris, sensitive_noise, on_matrices = "UD")
```

```{r}
noise[1:10, ]
```

## 3. Privacy

```{r}
# Re-identification rate
ri_rates = reidentification_rate(list("original" = iris, "anon" = swap), quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species"))

lapply(ri_rates, median)
```

- The re-identification rate has dropped significantly after the anonymization.

```{r}
prediction_plot(iris, k=5, swap)
```

## 4. Utility and similarity

- Also check the utility (performance on a model of interest, for example) and similarity (of means and variances, for example) of the anonymized data against those of the original data.

```{r, warning=FALSE}
# Utility
trainIndices = sample(c(TRUE,FALSE), nrow(iris), replace = TRUE, prob = c(0.7, 0.3))
train = iris[trainIndices, ]
test = iris[!trainIndices, ]
anon_train = spectral(train, sensitive_noise, on_matrices = "UD")

lm = lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species, data = train)
lm_anon = lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species, data = anon_train)
testRsquared_multimodel(list("Original" = lm, "Anon" = lm_anon), test, response = "Sepal.Length")
```

- The coefficient of determination on test data has dropped after the anonymization, meaning that the utility of the data has dropped.


```{r}
# Similarity
mediansAll(list("Original" = iris, "Anon" = swap))
```

- The means, variances and correlations have changed after the anonymization.
