---
title: "README"
output: github_document
---

# anon

A package containing data anonymization tools for R.

## 1. Installation

```{}
# 1. Install "devtools" package on R
    install.packages("devtools")

# 2. Install the "anon" package on R
    devtools::install_github("rajalah71/anon") # or /anon@paketti
```


```{r}
# 3. Load the "anon" library
    library(anon)
```

## 2. Usage

The package contains 4 functions to anonymize your data: $k$-anonymity, $l$-diversity, RSA encryption and spectral anonymization. 

The package also contains functions to measure the success of the anonymization on three levels: privacy, utility and similarity. 

### 2.1 _k_-anonymity

```{r}
data("iris")

kAnon(iris, k=5, quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species"))[1:10, ]
```




- This uses all columns as quasi identifiers by default and operates on them with mean on numeric columns and with mode on categorical ones.

- You can define your own generalization functions and on what columns the function will operate on. Functions must be on a named list which names match the names of the quasi identifier columns. The columns will be worked on the order you name them, or by cardinality if left unnamed. 

### 2.2 _l_-diversity

```{r}
lDiversity(iris, sensitiveAttributes = "Species", quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"), l=2)[1:10, ] 

```


- Works similiarly to kAnon, just need to specify the sensitive attributes. Won't probably work with more than one of them.

### 2.3 RSA

```{r}
encrypt(iris)[1:10, ]
```

- Encrypts the data using a new RSA key = 2048 bits by default. Do not distribute the public key.

### 2.4 Spectral anonymization

```{}
# General usage:
spectral(iris, your_function, on_matrices = "U")
```

- Anonymizes data on the spectral basis provided by SVD, $M = UDV'$. Supports any anoymization function $f: R^{a \times b} \to R^{a \times b}$.

```{r}
# Example usage, spectral permutation:
# Performs column permutations on each column the matrix U
data_anon = spectral(iris, cell_swap, on_matrices = "U")
data_anon[1:10, ]

# Example usage spectral noise:
# Adds random noise on the matrix columns of UD
spectral(iris, sensitive_noise, on_matrices = "UD")[1:10, ]

```

## 3. Privacy

```{r}
# Re-identification rate
ri_rates = reidentification_rate(list("original" = iris, "anon" = data_anon), quasiIdentifiers = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species"))

lapply(ri_rates, median)
```


```{r}
prediction_plot(iris, k=5, data_anon)
```


